---
title: "What made you happy today?"
output: html_notebook
---
Shengwei Huang 

## About the dataset

Many things can make one's heart smile with joy. HappyDB(https://rit-public.github.io/HappyDB/) is "a corpus of 100,000 crowd-sourced happy moments". The goal of this project is to look deeper into the causes that make us happy. Natural language processing and text mining are natural tools to derive interesting findings in this collection of happy moments.

## Part(0) Load and check packages 
```{r}

packages.used <- c("plyr","tm","tidytext","tidyverse","DT","wordcloud","scales","sentimentr","dplyr","ggplot2","topicmodels")  

# check whether we have dont have any of these packages 
packages.need <- setdiff(packages.used, intersect(installed.packages()[,1], packages.used))

# if we dont have that package/packages, install it/them.
if(length(packages.need)>0){
  install.packages(packages.need, dependencies = TRUE)
}
library(tm)
library(tidytext)
library(tidyverse)
library(DT)
library(wordcloud)
library(scales)
library(ggplot2)
library(sentimentr)
library(ngram)
library(dplyr)
library(topicmodels)
```

## Part (1) Data Preparation 

```{r}
# Here I will load two dataset: (1) the processed_moemnt.csv (2) demographic.csv
processed_moment <- read.csv("../output/processed_moments.csv")
demographic <- read.csv("../data/demographic.csv")
```


